---
output: github_document
editor_options: 
  chunk_output_type: console
---
<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r echo=FALSE}
knitr::opts_chunk$set(
  warning = FALSE, 
  message = FALSE,
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%",
  cache = TRUE
)
```
# Tailor

<!-- badges: start -->

```{r coverage, echo=FALSE}
cov <- covr::package_coverage()
pct <- covr::percent_coverage(cov)
pct_label <- sprintf("%.1f%%", pct)
  
# Choose badge color based on coverage
col <- if (pct >= 90) {
  "brightgreen"
  } else if (pct >= 75) {
    "yellow"
  } else if (pct >= 50) {
    "orange"
  } else {
    "red"
  }
  
badge_url <- sprintf(
  "https://img.shields.io/badge/coverage-%s-%s",
  URLencode(pct_label),
  col
)
```

[![status](https://www.repostatus.org/badges/latest/active.svg)](https://www.repostatus.org/#active)
[![R build status](https://github.com/sbthandras/tailor/workflows/R-CMD-check/badge.svg)](https://github.com/sbthandras/tailor/actions)
```{r coverage-badge, echo=FALSE, results='asis'}
cat(sprintf("[![Coverage](%s)](#test-coverage)\n\n", badge_url))
```
[![DOI](https://img.shields.io/badge/DOI-10.64898%2F2026.02.20.706991-blue)](https://doi.org/10.64898/2026.02.20.706991)

<!-- badges: end -->

`tailor` is a **T**ool for **A**dapter-domain **I**dentification and **L**inking **O**f **R**BPs.

## Installation

#### Install from Github (development version)

```{r install_github, eval=FALSE}
install.packages("devtools")
devtools::install_github("sbthandras/tailor")
```

## Usage

#### Identify shared adapter domain between two RBPs

```{r}
# load package
library(tailor)

# load example data set
data(rbps)

# align first two RBPs
ps <- position_scores(
  pattern = "MN395291-1", 
  subject = "ON513429-1", 
  id_var = "Core_ORF",
  seq_var = "translation",
  submat = "BLOSUM80",
  data = rbps
)

ps$position_scores |> head(5)

# find breakpoints along the alignment 
bps <- find_breakpoints(ps)

bps

# identify the conserved N-terminal domain if it exists
adapter <- find_adapter(bps)

adapter
```

Note, for positions 1-152 and 153-172 the scores were quite different, but both ranges were classified as conserved, so they were merged (default). To investigate this, let's plot the position scores:

```{r, out.width = '60%', dpi = 300}
plot(ps)
```

Depending on the method used for detecting breakpoints, the length of the identified adapter may differ. By default, `find_breakpoints()` uses `method = "cemean"` in which the function estimates the number of breakpoints and their locations using the Cross-Entropy Method from the `breakpoint` package. With `method = "plateau"` we get a different result:

```{r}
ps |> find_breakpoints(method = "plateau", type = "ewma") |> find_adapter()
```

Note that with this method we detect a shorter conserved N-terminal domain. 

#### Link RBPs by shared adapter domains

```{r, out.width = '60%', dpi = 300}
# load adapters of the example data set
data(adapters)

# convert to pident matrix and order by the original data frame
amat <- adapter_matrix(adapters, ids = rbps$Core_ORF)

# assign RBPs to clusters based on the distance matrix
clusters <- cluster_adapters(amat)

# visualise pident matrix with cluster assignments
plot(amat, clusters = clusters)
```

## Citation

If you use `tailor` in a publication, please cite our paper:

András Asbóth, Tamás Stirling, Orsolya Méhi, Gábor Apjok, Victor Klein-Sousa, Nicholas MI Taylor, Hiba Hadj Mehdi, Balázs Papp, Eszter Ari, Bálint Kintses (2026) A global map of receptor-binding protein compatibility for the programmable design of Klebsiella and Acinetobacter phages. https://doi.org/10.64898/2026.02.20.706991 

## Meta

- Please [report any issues, bugs or feature requests](https://github.com/sbthandras/tailor/issues).
- License: MIT
- Get citation information for `tailor` in R with `citation("tailor")`
